Assessing Power of the California Microhap Baseline for GSI (and calculating Fst)

### 2024-06-06 002-GSI_and_Fst.Rmd
Add summary of "recombinant" haplotypes
Remove spurious fixed allele diffs in scatters
Just an issue with calling NAs zero.  Fixed it.

@@

## Follow up on those Early alleles found in RRF and ERF

The analysis shows 4 fish from RRF and ERF with heterozygous genoypes (EL).
This has not been previously reported, so we really want to check up on those.
It might be that they had lots of missing data.

So, I want to grab their RoSA genotype strings so that we can verify.
```{r}
# here are the IDs fo the hets from those coastal pops
coastal_hets <- lfar_top_ass$top %>%
  left_join(rosa_genos, by = join_by(indiv)) %>%
  filter(collection %in% c("ERF", "RRF")) %>%
  select(indiv, rosa_geno) %>%
  filter(rosa_geno %in% c("EE", "EL"))
coastal_hets
```

Now, let's get the raw genotype strings (the "hapstrs") off those guys:
```{r}
rosa_hapstrs %>%
  semi_join(coastal_hets, by = join_by(indiv))
```

Those are definitely hets in the left half of the markers.  One of them is also
hets at the right half.  The three remaining ones are L homozygotes on the second
half.  So, there was been some recombination within these.

@@

# now get the freqs on there, order the collections and ensure
# that we are on the diagonal and the lower triangle, and then also get
# the positions for each of the points.  We will allow that the whole
# scatterplot can be inset into the cell by a fractional amount inset
inset <- 0.15
pair_freqs <- backbone %>%
left_join(alle_freqs, by = join_by(row_coll == collection, locus, allele)) %>%
rename(row_freq = freq) %>%
left_join(alle_freqs, by = join_by(col_coll == collection, locus, allele)) %>%
rename(col_freq = freq) %>%
mutate(
row_coll = factor(row_coll, levels = collection_order),
col_coll = factor(col_coll, levels = collection_order)
) %>%
filter(as.integer(col_coll) <= as.integer(row_coll)) %>%
replace_na(list(row_freq = 0.0, col_freq = 0.0)) %>%
mutate( # this is just setting things as in table_as_figure
ymin = nlevels(row_coll) - as.integer(row_coll) + inset,
ymax = nlevels(row_coll) - as.integer(row_coll) + 1 - inset,
xmin = as.integer(col_coll) - 1 + inset,
xmax = as.integer(col_coll) - inset
) %>%
mutate(  # here we get the x-y positions for the scatter points
x = xmin + col_freq * (xmax - xmin),
y = ymin + row_freq * (ymax - ymin)
)

>>>>

## Let's do a quick summary of the apparent recombinants

```{r, rows.print=30}
H8_recombs %>%
  left_join(
    full_base2 %>% select(indiv, repunit, collection),
    by = join_by(indiv)
  ) %>% 
  arrange(hapstr8)
```

@@

## Look at the distribution of the E2 alleles across collections

@@

# now get the freqs on there, order the collections and ensure
# that we are on the diagonal and the lower triangle, and then also get
# the positions for each of the points.  We will allow that the whole
# scatterplot can be inset into the cell by a fractional amount inset
inset <- 0.15
pf_init <- backbone %>%
  left_join(alle_freqs, by = join_by(row_coll == collection, locus, allele)) %>%
  rename(row_freq = freq) %>%
  left_join(alle_freqs, by = join_by(col_coll == collection, locus, allele)) %>%
  rename(col_freq = freq) %>%
  mutate(
    row_coll = factor(row_coll, levels = collection_order),
    col_coll = factor(col_coll, levels = collection_order)
  ) %>%
  filter(as.integer(col_coll) <= as.integer(row_coll))

# now, here is a somewhat more complex step.  The problem is that some collections have
# no observations of any loci, it seems.  In those cases, we want the NA values to stand.
# Those cases can be recognized because every allele at a locus in them is NA. In cases,
# where not all of the alleles have NA for frequency, we do want to set the NAs to 0.
pf0s <- pf_init %>%
  group_by(row_coll, col_coll, locus) %>%
  mutate(
    allNA_rowfreq = all(is.na(row_freq)),
    allNA_colfreq = all(is.na(col_freq)),
  ) %>%
  ungroup() %>%
  mutate(
    row_freq = ifelse(is.na(row_freq) & !allNA_rowfreq, 0.0, row_freq),
    col_freq = ifelse(is.na(col_freq) & !allNA_colfreq, 0.0, col_freq)
  )

# finally, do some more setting up of things:
pair_freqs <- pf0s %>% 
  mutate( # this is just setting things as in table_as_figure
    ymin = nlevels(row_coll) - as.integer(row_coll) + inset,
    ymax = nlevels(row_coll) - as.integer(row_coll) + 1 - inset,
    xmin = as.integer(col_coll) - 1 + inset,
    xmax = as.integer(col_coll) - inset
  ) %>%
  mutate(  # here we get the x-y positions for the scatter points
    x = xmin + col_freq * (xmax - xmin),
    y = ymin + row_freq * (ymax - ymin)
  )
